On the validity of the Born approximation for beyond-Gaussian weak lensing observables

### 2017-01-25 Paper/paper.tex
finished modifications for resubmission (2)

\begin{abstract}
Accurate forward modeling of weak lensing (WL) observables from cosmological parameters is necessary for upcoming galaxy surveys. Because WL probes structures in the non--linear regime, analytical forward modeling is very challenging, if not impossible. Numerical simulations of WL features rely on ray--tracing through the outputs of $N$--body simulations, which requires knowledge of the gravitational potential and accurate solvers for light ray trajectories. A less accurate procedure, based on the Born approximation, only requires knowledge of the density field, and can be implemented more efficiently and at a lower computational cost. \textbf{\color{red} In this work, we use simulations to show that deviations of the Born--approximated convergence power spectrum, skewness and kurtosis from their fully ray--traced counterparts are consistent with the smallest non--trivial $O(\Phi^3)$ post--Born corrections (so-called geodesic and lens-lens terms).  In this work, we use simulations to show that deviations of the Born--approximated convergence power spectrum, skewness and kurtosis from their fully ray--traced counterparts are consistent with the smallest non--trivial post--Born corrections. Our results imply a cancelation among the larger $O(\Phi^4)$ (and higher order) terms, consistent with previous analytic work.} We also find that cosmological parameter biases induced by the Born approximation are negligible even for an LSST--like survey, once galaxy shape noise is considered, and we conclude that the Born approximation is sufficient for future galaxy WL analyses. Using the \ttt{LensTools} software suite, we show that the Born approximation saves a factor of 4 in computing time with respect to the full ray--tracing in reconstructing the convergence.       
\end{abstract}

@@

\begin{figure*}
\begin{center}
\includegraphics[scale=0.46]{Figures/delta_m2.eps}\includegraphics[scale=0.46]{Figures/delta_m5.eps}
\end{center}
\caption{Dominant post--Born corrections to the $\kappa$ skewness and kurtosis due to lens--lens coupling and geodesic perturbation, as a function of the scale $\theta_G$ of the Gaussian smoothing window. We show the contributions of the lens--lens couplings (red) and the geodesic perturbation (green). We show both the cases in which the residuals $\delta\kappa^3,\delta\kappa^4$ are computed adding equations (\ref{sim:ll},\ref{sim:gp}) to the Born field (solid lines) and in which the residuals are computed from the dominant $\Phi$ orders only as in equations (\ref{res:skewcross},\ref{res:kurtcross}) (dashed lines). \textbf{\color{red} For reference we show the first non--trivial corrections due to reduced shear (black lines).} The quantities shown are the ensemble averages over 8192 realizations of $\kappa$ in the fiducial cosmology $\bb{p}_0$.}
\label{fig:skResiduals}
\end{figure*}

@@

\section{Discussion}
\label{sec:discuss}
% 
In this section, we discuss our main findings. Figure \ref{fig:csample} shows that the dominant post--Born correction to the $\kappa$ reconstruction comes from the geodesic term (eq.~\ref{sim:gp}). This can be seen by looking at the lens-lens term, which is over an order of magnitude smaller than the geodesic term, and by looking at the full $\kappa-\kappa_{\rm born}$ residuals, which show a very similar overall level, as well as structural detail, to the ones found in the $\kappa_{\rm geo}$ map. \textbf{\color{red} When considering a real observation, in which $\kappa$ is estimated from source ellipticities, one should consider an additional reduced shear correction to $\kappa$, which is of order $O(\Phi^2)$ \citep{HirataKrause} and is proportional to $\gamma\kappa$. We estimated this correction to be one order of magnitude smaller than the dominant geodesic term}. The reader might have noticed the curious alignment feature in Figure \ref{fig:csample}, in which all the dipolar structures in $\kappa_{\rm geo}$ appear to be horizontally aligned throughout the field of view. We believe that this alignment is not a numerical effect, but is originated physically by the $\nabla\delta$ terms in eq.~(\ref{sim:gp}). Early in the line--of--sight integration, the field of view covers a very small section of the simulation box ($\sim7\,{\rm Mpc}$ for the first lens), within which dipolar structures are aligned. These alignments survive after the full integration is completed due to the $\nabla\delta\cdot\nabla\Phi$ term, which couples lenses at different redshifts. To provide evidence that the closest lenses to the observer are responsible for the alignment features, we performed the line--of--sight integration for $\kappa_{\rm geo}$ by rotating the first 5 lenses by 90 degrees. This is sufficient to coherently flip the dipolar structures from horizontal to vertical throughout the field of view.

Figure \ref{fig:psResiduals} shows a comparison between the Born--approximated $\kappa$ power spectra and the ones obtained with full ray--tracing, separating different orders in $\Phi$. We can conclude that the dominant contributions to the residuals come from the Born--geodesic cross power, which dominates the Born--lens--lens cross power by two orders of magnitude on small scales. Figure \ref{fig:skResiduals} shows that the geodesic and lens--lens contributions to the $\kappa$ skewness and kurtosis are comparable and that the lowest non--trivial post--Born corrections can fully account for the discrepancy between the Born--approximated quantities and the fully ray--traced ones. Figs. \ref{fig:psResiduals} and \ref{fig:skResiduals} indicate \textbf{\color{red} that truncating $\kappa$ at $O(\Phi^2)$ post--Born terms, leads to inaccurate predictions of the power spectrum and moments. In particular, higher-order terms involving $\langle \kappa_{\rm born}^n \kappa_{\rm ll}^m \kappa_{\rm geo}^2\rangle$ and higher, cause large deviations (as seen by the difference between the solid and dashed green curves) that, in an exact approach, are canceled by $O(\Phi^3)$ corrections to $\kappa$, as pointed out by \citep{HirataKrause,CMBPrattenLewis,ScoccimarroLoop}}. Therefore, in practice, in order to go beyond Born approximation, it is more advantageous, both for accuracy and for CPU time, to perform full ray--tracing (whose computational cost is comparable to those of the $O(\Phi^2)$ line--of--sight integrals; Table \ref{tab:benchmarks}).

>>>>

\begin{abstract}
Accurate forward modeling of weak lensing (WL) observables from cosmological parameters is necessary for upcoming galaxy surveys. Because WL probes structures in the non--linear regime, analytical forward modeling is very challenging, if not impossible. Numerical simulations of WL features rely on ray--tracing through the outputs of $N$--body simulations, which requires knowledge of the gravitational potential and accurate solvers for light ray trajectories. A less accurate procedure, based on the Born approximation, only requires knowledge of the density field, and can be implemented more efficiently and at a lower computational cost. \textbf{\color{red} In this work, we use simulations to show that deviations of the Born--approximated convergence power spectrum, skewness and kurtosis from their fully ray--traced counterparts are consistent with the smallest non--trivial $O(\Phi^3)$ post--Born corrections (so-called geodesic and lens-lens terms). Our results imply a cancelation among the larger $O(\Phi^4)$ (and higher order) terms, consistent with previous analytic work.} We also find that cosmological parameter biases induced by the Born approximation are negligible even for an LSST--like survey, once galaxy shape noise is considered, and we conclude that the Born approximation is sufficient for future galaxy WL analyses. Using the \ttt{LensTools} software suite, we show that the Born approximation saves a factor of 4 in computing time with respect to the full ray--tracing in reconstructing the convergence.       
\end{abstract}

@@

\begin{figure*}
\begin{center}
\includegraphics[scale=0.46]{Figures/delta_m2.eps}\includegraphics[scale=0.46]{Figures/delta_m5.eps}
\end{center}
\caption{Dominant post--Born corrections to the $\kappa$ skewness and kurtosis due to lens--lens coupling and geodesic perturbation, as a function of the scale $\theta_G$ of the Gaussian smoothing window. We show the contributions of the lens--lens couplings (red) and the geodesic perturbation (green). We show both the cases in which the residuals $\delta\kappa^3,\delta\kappa^4$ are computed adding equations (\ref{sim:ll},\ref{sim:gp}) to the Born field (solid lines) and in which the residuals are computed from the dominant $\Phi$ orders only as in equations (\ref{res:skewcross},\ref{res:kurtcross}) (dashed lines). \textbf{\color{red} For reference we show the first non--trivial corrections due to reduced shear (black lines), which can be calculated from the Born images.} The quantities shown are the ensemble averages over 8192 realizations of $\kappa$ in the fiducial cosmology $\bb{p}_0$.}
\label{fig:skResiduals}
\end{figure*}

@@

\section{Discussion}
\label{sec:discuss}
% 
In this section, we discuss our main findings. Figure \ref{fig:csample} shows that the dominant post--Born correction to the $\kappa$ reconstruction comes from the geodesic term (eq.~\ref{sim:gp}). This can be seen by looking at the lens-lens term, which is over an order of magnitude smaller than the geodesic term, and by looking at the full $\kappa-\kappa_{\rm born}$ residuals, which show a very similar overall level, as well as structural detail, to the ones found in the $\kappa_{\rm geo}$ map. The reader might have noticed the curious alignment feature in Figure \ref{fig:csample}, in which all the dipolar structures in $\kappa_{\rm geo}$ appear to be horizontally aligned throughout the field of view. We believe that this alignment is not a numerical effect, but is originated physically by the $\nabla\delta$ terms in eq.~(\ref{sim:gp}). Early in the line--of--sight integration, the field of view covers a very small section of the simulation box ($\sim7\,{\rm Mpc}$ for the first lens), within which dipolar structures are aligned. These alignments survive after the full integration is completed due to the $\nabla\delta\cdot\nabla\Phi$ term, which couples lenses at different redshifts. To provide evidence that the closest lenses to the observer are responsible for the alignment features, we performed the line--of--sight integration for $\kappa_{\rm geo}$ by rotating the first 5 lenses by 90 degrees. This is sufficient to coherently flip the dipolar structures from horizontal to vertical throughout the field of view.

Figure \ref{fig:psResiduals} shows a comparison between the Born--approximated $\kappa$ power spectra and the ones obtained with full ray--tracing, separating different orders in $\Phi$. We can conclude that the dominant contributions to the residuals come from the Born--geodesic cross power, which dominates the Born--lens--lens cross power by two orders of magnitude on small scales. Figure \ref{fig:skResiduals} shows that the geodesic and lens--lens contributions to the $\kappa$ skewness and kurtosis are comparable and that the lowest non--trivial post--Born corrections can fully account for the discrepancy between the Born--approximated quantities and the fully ray--traced ones. \textbf{\color{red} The geodesic and lens--lens corrections to the skewness we find agree, as orders of magnitude, with the ones already published in \citep{WLBispectrumDodelson}, although our measured reduced shear correction is larger. We stress, however, that our estimates and the ones in \citep{WLBispectrumDodelson} have been produced under different assumptions, in particular \citep{WLBispectrumDodelson} use a fully analytical approach based on the Limber approximation, different form forms for the smoothing windows and use their own assumptions for the form of the $\kappa$ bispectrum.} Figs. \ref{fig:psResiduals} and \ref{fig:skResiduals} indicate \textbf{\color{red} that truncating $\kappa$ at $O(\Phi^2)$ post--Born terms, leads to inaccurate predictions of the power spectrum and moments. In particular, higher-order terms involving $\langle \kappa_{\rm born}^n \kappa_{\rm ll}^m \kappa_{\rm geo}^2\rangle$ and higher, cause large deviations (as seen by the difference between the solid and dashed green curves) that, in an exact approach, are canceled by $O(\Phi^3)$ corrections to $\kappa$, as pointed out by \citep{HirataKrause,CMBPrattenLewis,ScoccimarroLoop}}. Therefore, in practice, in order to go beyond Born approximation, it is more advantageous, both for accuracy and for CPU time, to perform full ray--tracing (whose computational cost is comparable to those of the $O(\Phi^2)$ line--of--sight integrals; Table \ref{tab:benchmarks}).
